KidsAIR: Main analysis initial steps

### 2020-09-30 main_analysis_initial.Rmd
analysis updates

---
title: "KidsAIR: Main analysis initial steps"
author: "Ethan Walker"
date: "Started 23 Sept 2020, Updated 28 Sept 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

@@

```{r}
# Initial summary data for LRTI
incidence_summary <- analysis_data %>% 
  group_by(treatment) %>% 
  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
            sd_inc = sd(lrti_incidence, na.rm = TRUE),
            min_inc = min(lrti_incidence, na.rm = TRUE),
            median_inc = median(lrti_incidence, na.rm = TRUE),
            max_inc = max(lrti_incidence, na.rm = TRUE))
incidence_summary
```

```{r}
model_results <- glmer(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data, family = poisson(link = "log"))

>>>>

---
title: "KidsAIR: Main analysis initial steps"
author: "Ethan Walker"
date: "Started 23 Sept 2020, Updated 30 Sept 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

@@

```{r}
# clean up data for table 1 summary
table_1_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  distinct(area, home, .keep_all = TRUE)

# summary data for manuscript tables
char_funct_total <- function(var, data = table_1_data){

demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(total_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = round(n()/total_n*100)) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_treatment <- function(var, data = table_1_data){

demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  group_by(treatment) %>% 
  mutate(treatment_n = n()) %>% 
  group_by(treatment, new_var) %>% 
    mutate(n = n(),
         percent = round(n()/treatment_n*100)) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  select(new_var, n, percent) %>% 
  arrange(treatment, new_var)
demographics_summary
}

char_funct_total("gender_parent")
char_funct_treatment("gender_parent")
char_funct_total("hispanic_parent")
char_funct_treatment("hispanic_parent")
char_funct_total("race_parent")
char_funct_treatment("race_parent")
char_funct_total("education_3level")
char_funct_treatment("education_3level")
char_funct_total("income_3level")
char_funct_treatment("income_3level")
char_funct_total("home_floors_2level")
char_funct_treatment("home_floors_2level")
char_funct_total("home_year_built_2level")
char_funct_treatment("home_year_built_2level")
char_funct_total("home_pets_2level")
char_funct_treatment("home_pets_2level")
char_funct_total("home_pets")
char_funct_treatment("home_pets")
char_funct_total("stove_age_3level")
char_funct_treatment("stove_age_3level")
char_funct_total("chimney_clean_3level")
char_funct_treatment("chimney_clean_3level")
char_funct_total("wood_collect_method_2level")
char_funct_treatment("wood_collect_method_2level")
char_funct_total("wood_collect_2level")
char_funct_treatment("wood_collect_2level")
char_funct_total("stove_grade_3level")
char_funct_treatment("stove_grade_3level")
char_funct_total("burn_level_3level")
char_funct_treatment("burn_level_3level")
char_funct_total("ruca_code")
char_funct_treatment("ruca_code")
char_funct_total("ruca_code_2")
char_funct_treatment("ruca_code_2")
char_funct_total("ruca_code_3")
char_funct_treatment("ruca_code_3")
```

```{r}
# clean up data for table 1 summary
table_1_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  distinct(area, home, .keep_all = TRUE)

# summary data for manuscript tables
num_funct_total <- function(var, data = table_1_data){

demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  filter(!is.na(new_var)) %>% 
  select(treatment, home, new_var) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_treatment <- function(var, data = table_1_data){

demographics_summary <- data %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
  filter(!is.na(new_var)) %>% 
  select(treatment, home, new_var) %>% 
  group_by(treatment) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_total("residents_under_five")
num_funct_treatment("residents_under_five")
num_funct_total("home_bedrooms")
num_funct_treatment("home_bedrooms")
num_funct_total("pm_mean_sampling_period")
num_funct_treatment("pm_mean_sampling_period")
num_funct_total("pm_sample_interval")
num_funct_treatment("pm_sample_interval")
num_funct_total("moisture_closest")
num_funct_treatment("moisture_closest")
num_funct_total("temp_indoor_max")
num_funct_treatment("temp_indoor_max")
num_funct_total("rh_indoor_max")
num_funct_treatment("rh_indoor_max")
num_funct_total("mean_temp")
num_funct_treatment("mean_temp")
num_funct_total("amb_pm_24hr")
num_funct_treatment("amb_pm_24hr")
num_funct_total("sums_events_sampling_period_2.5")
num_funct_treatment("sums_events_sampling_period_2.5")
num_funct_total("sums_events_sampling_period_5")
num_funct_treatment("sums_events_sampling_period_5")
num_funct_total("sums_events_sampling_period_7.5")
num_funct_treatment("sums_events_sampling_period_7.5")
```

```{r}
# Initial summary data for LRTI
incidence_summary <- analysis_data %>% 
  group_by(treatment) %>% 
  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
            sd_inc = sd(lrti_incidence, na.rm = TRUE),
            n_inc = n(),
            min_inc = min(lrti_incidence, na.rm = TRUE),
            median_inc = median(lrti_incidence, na.rm = TRUE),
            max_inc = max(lrti_incidence, na.rm = TRUE))
incidence_summary

# Histogram for distribution of data
incidence_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_incidence)) +
    theme_light() +
    facet_wrap(~treatment)
incidence_plot

# Check for overdispersion in data using likelihood ratio test
library(pscl)
library(MASS)

nb_results <- glm.nb(lrti_incidence ~ treatment, data = analysis_data)
odTest(nb_results)

# According to these results, data is overdispersed
# Variance is much higher than the mean for LRTI incidence
# Assumption of Poisson is variance = mean, so recommend using negative binomial model

summary(nb_results)

# Link for analysis ideas and notes:
# https://stats.idre.ucla.edu/r/dae/negative-binomial-regression/ 
# https://stats.idre.ucla.edu/r/dae/poisson-regression/ 
# https://stats.idre.ucla.edu/r/faq/random-coefficient-poisson-models/ 
```

```{r}
# Mixed model
model_results <- glmer.nb(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data)
